The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-15 script/en/chapters/paper_and_pencil.tex
CT script en: Sage examples on working with the shift and Caesar ciphers using the affine cipher

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: shiftcipher = ShiftCryptosystem(AlphabeticStrings()); shiftcipher
Shift cryptosystem on Free alphabetic string monoid on A-Z
sage: plaintext = shiftcipher.encoding("Shift me any number of positions.")
sage: plaintext
SHIFTMEANYNUMBEROFPOSITIONS
sage: key = 12  # shift by 12 positions
sage: 
sage: # shift the plaintext by 12 positions to get the ciphertext
sage: ciphertext = shiftcipher.enciphering(key, plaintext)
sage: ciphertext
ETURFYQMZKZGYNQDARBAEUFUAZE
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = shiftcipher.deciphering(key, ciphertext); P
SHIFTMEANYNUMBEROFPOSITIONS
sage: P == plaintext
True
\end{Verbatim}
\caption{A shift cipher over the upper-case letters of the English alphabet.}
\end{sagecode}

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # create a Caesar cipher
sage: caesarcipher = ShiftCryptosystem(AlphabeticStrings())
sage: plaintext = caesarcipher.encoding("Shift the alphabet by three positions to the right.")
sage: plaintext
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: 
sage: # shift the plaintext by 3 positions
sage: key = 3
sage: ciphertext = caesarcipher.enciphering(key, plaintext)
sage: ciphertext
VKLIWWKHDOSKDEHWEBWKUHHSRVLWLRQVWRWKHULJKW
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = caesarcipher.deciphering(key, ciphertext)
sage: P
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: P == plaintext
True
\end{Verbatim}
\caption{Constructing the Caesar cipher using the shift cipher.}
\end{sagecode}

@@

>>>>

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: shiftcipher = ShiftCryptosystem(AlphabeticStrings()); shiftcipher
Shift cryptosystem on Free alphabetic string monoid on A-Z
sage: plaintext = shiftcipher.encoding("Shift me any number of positions.")
sage: plaintext
SHIFTMEANYNUMBEROFPOSITIONS
sage: key = 12  # shift by 12 positions
sage: 
sage: # shift the plaintext by 12 positions to get the ciphertext
sage: ciphertext = shiftcipher.enciphering(key, plaintext)
sage: ciphertext
ETURFYQMZKZGYNQDARBAEUFUAZE
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = shiftcipher.deciphering(key, ciphertext); P
SHIFTMEANYNUMBEROFPOSITIONS
sage: P == plaintext
True
\end{Verbatim}
\caption{A shift cipher over the upper-case letters of the English alphabet.}
\label{paper_pencil:shift_cipher:Sage_example}
\end{sagecode}

@@

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # create a Caesar cipher
sage: caesarcipher = ShiftCryptosystem(AlphabeticStrings())
sage: plaintext = caesarcipher.encoding("Shift the alphabet by three positions to the right.")
sage: plaintext
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: 
sage: # shift the plaintext by 3 positions
sage: key = 3
sage: ciphertext = caesarcipher.enciphering(key, plaintext)
sage: ciphertext
VKLIWWKHDOSKDEHWEBWKUHHSRVLWLRQVWRWKHULJKW
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = caesarcipher.deciphering(key, ciphertext)
sage: P
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: P == plaintext
True
\end{Verbatim}
\caption{Constructing the Caesar cipher using the shift cipher.}
\label{paper_pencil:shift_cipher:Sage_example_Caesar_cipher}
\end{sagecode}

@@

We can also construct a shift cipher using the affine cipher. To do
so, we need to restrict keys of the affine cipher be of the form 
$(1, b)$ where $b$ is any non-negative integer. For instance, we can
work through example~\ref{paper_pencil:shift_cipher:Sage_example} on
page~\pageref{paper_pencil:shift_cipher:Sage_example} as follows:

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # construct an affine cipher
sage: shiftcipher = AffineCryptosystem(AlphabeticStrings()); shiftcipher
Affine cryptosystem on Free alphabetic string monoid on A-Z
sage: plaintext = shiftcipher.encoding("Shift me any number of positions.")
sage: plaintext
SHIFTMEANYNUMBEROFPOSITIONS
sage: 
sage: # shift the plaintext by 12 positions to get the ciphertext
sage: a, b = (1, 12)
sage: ciphertext = shiftcipher.enciphering(a, b, plaintext)
sage: ciphertext
ETURFYQMZKZGYNQDARBAEUFUAZE
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = shiftcipher.deciphering(a, b, ciphertext); P
SHIFTMEANYNUMBEROFPOSITIONS
sage: P == plaintext
True
\end{Verbatim}
\caption{Constructing a shift cipher using the affine cipher.}
\end{sagecode}

We can also use the affine cipher to create the Caesar cipher. To do
so, the encryption/decryption key must be $(1, 3)$. In the next
example, we work through
example~\ref{paper_pencil:shift_cipher:Sage_example_Caesar_cipher} on
page~\pageref{paper_pencil:shift_cipher:Sage_example_Caesar_cipher}
using the affine cipher.

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # create a Caesar cipher
sage: caesarcipher = AffineCryptosystem(AlphabeticStrings())
sage: plaintext = caesarcipher.encoding("Shift the alphabet by three positions to the right.")
sage: plaintext
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: 
sage: # shift the plaintext by 3 positions
sage: a, b = (1, 3)
sage: ciphertext = caesarcipher.enciphering(a, b, plaintext)
sage: ciphertext
VKLIWWKHDOSKDEHWEBWKUHHSRVLWLRQVWRWKHULJKW
sage: 
sage: # decrypt the ciphertext and ensure that it is the original plaintext
sage: P = caesarcipher.deciphering(a, b, ciphertext)
sage: P
SHIFTTHEALPHABETBYTHREEPOSITIONSTOTHERIGHT
sage: P == plaintext
True
\end{Verbatim}
\caption{Constructing the Caesar cipher using the affine cipher.}
\end{sagecode}
